### 2023-05-03 manuscript.tex
Draft remaining parts
* manuscript.tex: Add remaining chapters; update title page.
* reference.bib: Add referenced entries.

\onehalfspacing
\setlist{noitemsep}
\titleformat{\chapter}{\normalfont\huge\bfseries}{\thechapter}{1em}{}
\titlespacing*{\chapter}{
  0pt}{3.5ex plus 1ex minus .2ex}{2.3ex plus .2ex}

@@

\newcommand{\context}{\mathrel{/}}
\newcommand{\cuhk}{The Chinese University of Hong Kong}
\newcommand{\gap}{\underline{\hspace{1em}}}
\newcommand{\textemph}[1]{\textsc{#1}}
\newcommand{\textphon}[1]{[#1]}
\newcommand{\textterm}[1]{\textsc{#1}\index{#1}}

\renewcommand{\implies}{\Rightarrow}
\renewcommand{\ExLBr}{(\thechapter.}

@@

    A Thesis To Be Submitted in Partial Fulfilment of

@@

    \monthyyyy\formatdate{28}{12}{2022}
  \end{center}
  \pretocmd{\doclicenseLongText}{%
    Copyright \copyright\ 2023 Wing Hei Chan\par}{}{}
  \doclicenseThis
  \vspace*{\fill}
\end{titlepage}

@@

>>>>

\onehalfspacing
\setlength{\jot}{0pt}
\setlist{noitemsep}
\titleformat{\chapter}{\normalfont\huge\bfseries}{\thechapter}{1em}{}
\titlespacing*{\chapter}{
  0pt}{3.5ex plus 1ex minus .2ex}{2.3ex plus .2ex}

@@

\newcommand{\context}{\mathrel{/}}
\newcommand{\cuhk}{The Chinese University of Hong Kong}
\newcommand{\gap}{\underline{\hspace{1em}}}
\newcommand{\shift}{\hspace*{1em}}
\newcommand{\textemph}[1]{\textsc{#1}}
\newcommand{\textfeat}[1]{\textsc{#1}}
\newcommand{\textfor}[1]{\textit{#1}}
\newcommand{\textform}[1]{\textit{#1}}
\newcommand{\textgls}[1]{`#1'}
\newcommand{\textphon}[1]{[#1]}
\newcommand{\textterm}[1]{\textsc{#1}\index{#1}}

\renewcommand{\implies}{\Rightarrow}
\renewcommand{\ExLBr}{(\thechapter.}
\renewenvironment{flushleft}{\raggedright}{}

@@

    A Thesis Submitted in Partial Fulfilment of

@@

    \monthyyyy\formatdate{08}{05}{2023}
  \end{center}
  \pretocmd{\doclicenseLongText}{%
    Copyright \copyright\ 2023 Wing Hei Chan\par}{}{}
  \doclicenseThis
  \vspace*{\fill}
\end{titlepage}

@@

\chapter{Boundary of Transformation}
With the computational preliminaries in place, we are now ready to
analyze the attested tonal transformations found in Chinese languages.
This chapter specifically discusses transformations involving
boundaries that can block the applications.  Despite the simple facets
of transformations \textfor{per se}, the existence of boundaries
raises certain important questions, such as where, when, and how they
come into play.  This is especially problematic under linear models,
as hierarchical interpretations of boundaries often require more than
linear sequences.

\section{Representation of Boundary}
The linear understanding of \textterm{boundaries} is simply certain
terms that never surface, yet block any transformation that may
otherwise apply.  A boundary term is always output \textemph{as is},
therefore failing to satisfy any condition when it precedes or
succeeds the input term in question.  Consider again the unbounded
tone spreading rule:

\ex. \(\acute{\sigma}'(x) \coloneq \acute{\sigma}(x) \lor
(\sigma(x) \land \acute{\sigma}'(\mathit{pred}(x)))\)

A boundary term certainly cannot satisfy \(\sigma(x)\), and thus
\(\acute{\sigma}'(\mathit{pred}(x))\) cannot be true when it precedes
the input term.  This quickly fails to account for more complicated
rule interactions, where boundaries may \textemph{change} to create
new contexts.  Recall that cyclicity assumes that each cycle creates
new contexts until fixed point---this intuition leads to analyses that
delete and insert new boundaries.

An alternative view of boundaries is \textterm{groupings}, which can
render the transformation nonregular.  Even if we ignore this nature
of arbitrary groupings by restricting self-embedding in some manner,
there is still the issue of determining whether two terms are in the
same group.  Suppose groupings are delimited by conventional brackets,
to determine whether two terms are in the same group, the transducer
potentially needs to traverse the whole input sequence.  This is more
apparent in the case of autosegmental representations, where groupings
are represented as represented as trees, which are a subset of
directed graphs.\footnote{Such trees with a limited depth are known as
  \enquote{prosodic hierarchy}, for which see \textcite{nv86pp}.} The
relation that decides whether two terms are in the same group is then

\ex. \(\mathit{SameGroup}(x, y) \coloneq \exists z\ldotp
\mathit{Assoc}(z, x) \land \mathit{Assoc}(z, y)\)

assuming that groupings do not nest.  As usual, nestable groupings
require the transitive closure of \(\mathit{Assoc}\), which is monadic
second-order definable.

If we entirely discard the idea of explicitly encoding boundaries in
input sequences, we need to derive the same processes as before,
thereby implicitly encoding boundaries in some sense.  What boundary
terms achieve is essentially applying a single transformation to
multiple linear sequences and combining the outputs through some
associative operation, say concatenation.  In other words, the
transformation is \enquote{lifted} to operate on multiple linear
sequences.  By doing so, we modularize the transformation into a
simpler transformation and a universal lift operation.\footnote{This
  operation is precisely a functor, or loosely speaking function of
  type
  \((\alpha \to \beta) \to \mathbb{F}\alpha \to \mathbb{F}\beta\),
  commonly known as \(\mathit{map}\).  \(\mathbb{F}\) should produce a
  type whose values can be \enquote{folded} in the sense of
  \textcite{h99tuem} under a monoid.}

However, groupings are much different, as they create input
\textemph{trees}.  Input trees, by their inductive nature, can contain
more trees, which are in turn input trees to a potentially new
transducer with appropriate hierarchical notions encoded.  It not only
generates a recursive process, but also creates a leaky abstraction as
the transducer must be aware of the whole input up to a certain level.
This defeats the purpose of having multiple linguistic levels in the
first place.

The ideal scenario is that each linguistic level drives phonological
realization by feeding the next level inputs and combining the
outputs.  This eliminates the need of boundaries altogether, while
implicitly preserving their essence.  This is also much what cyclicity
really aims to achieve, namely that each cycle further transforms the
input sequence with one layer of groupings removed.  What this means
in practice will be demonstrated in the following sections as relevant
linguistic data are covered.

\section{Standard Chinese: Tone 3 Sandhi}
Tone 3 sandhi is a tonal transformation found in Standard Chinese,
where tone 3 is dissimilated to tone 2 before another tone 3.  This
process can be described as simultaneously applied, as seen in the
following data:

\ex.
\a. \(\text{\textform{zhǎnlǎn}} \to
\text{\textform{zhánlǎn} \textgls{exhibit}}\)
\b. \(\text{\textform{zhǎnlǎnguǎn}} \to
\text{\textform{zhánlánguǎn} \textgls{exhibition hall}}\)
\b. \(\text{\textform{zhǎnlǎnguǎn lǐ}} \to
\text{\textform{zhánlánguán lǐ} \textgls{in exhibition hall}}\)

Therefore, a logical transduction

\ex. \(\acute{\sigma}'(x) \coloneq \acute{\sigma}(x) \lor
(\check{\sigma}(x) \land \check{\sigma}(\mathit{succ}(x)))\)

resulting in the mapping

\ex. \(\check{\sigma}^{n+1} \to \acute{\sigma}^{n}\check{\sigma}\)

should be enough to describe the process.  It should be emphasized
that this process is linear in its computational nature,
\textemph{regardless} of the specific representation used here for
convenience.  Indeed, even if the tones are represented on a separate
tier with multiple terms, the transformation is still linear and,
importantly, strictly local, albeit with a larger memory
size.\footnote{To decide the memory size, it suffices to calculate the
  number \(n + m\), where \(n\) and \(m\) are respectively the highest
  degrees of composition of the predecessor and successor functions in
  the normalized transduction.}  Namely, the mapping will then be

\ex. \((\text{LL})^{n+1} \to (\text{LR})^{n}\text{LL}\)

ignoring the optional alternative realization of tone 3 when it is
lengthened \parencite{d99mstems}.  The fundamental reason is that the
associations are not affected by the transformation, that is,
\(\mathit{Assoc}\) is never mentioned in the transduction.

The transformation becomes more interesting when boundaries are
concerned.  The data

\ex.
\a. \(\text{\textform{gǒu bǐ mǎ xiǎo}} \to
\text{\textform{góu bǐ má xiǎo}}\)
\bg. gǒu bǐ mǎ xiǎo\\
dog than horse small\\
\trans\textgls{Dogs are smaller than horses.}

show that certain tone 3 does \textemph{not} change to tone 2.  This
seems to suggest that an alternative transduction

\ex. \(\acute{\sigma}'(x) \coloneq \acute{\sigma}(x) \lor
(\check{\sigma}(x) \land \check{\sigma}'(\mathit{succ}(x)))\)

resulting in the mapping

\ex.
\a. \(\check{\sigma}^{2n} \to (\acute{\sigma}\check{\sigma})^{n}\)
\b. \(\check{\sigma}^{2n+1} \to
\check{\sigma} (\acute{\sigma}\check{\sigma})^{n}\)

is in effect.  This is not true, as the data

\ex.
\a. \(\text{\textform{xiǎogǒu bǐ xiǎomǎ xiǎo}} \to
\text{\textform{xiáogóu bǐ xiáomá xiǎo}}\)
\bg. xiǎo-gǒu bǐ xiǎo-mǎ xiǎo\\
small-dog than small-horse small\\
\trans\textgls{Puppies are smaller than ponies.}

clearly do not adhere to the mapping.  Moreover, the data

\ex.
\a. \(\text{\textform{xiǎo zhǐlǎohǔ}} \to
\{\text{\textform{xiǎo zhíláohǔ}},
\text{\textform{xiáo zhǐláohǔ}},
\text{\textform{xiáo zhíláohǔ}}\}\)
\bg. xiǎo zhǐ-lǎohǔ\\
small paper-tiger\\
\trans\textgls{small paper tiger}

show that certain tone 3 \textemph{optionally} changes to tone 2.  To
account for such contrasts, various analyses have been developed.  Two
such analyses, where one is representative for \enquote{syntactic}
analyses, and the other for \enquote{prosodic} analyses, will be
discussed.

\subsection{Syntactic Analysis}
\textcite{c00tspcd} argues for a analysis under which feet are built
based on syntactic structures, which in turn govern the application of
tone 3 sandhi.  As \textcite{d07psc} states, Although this analysis
includes a notion of feet, they are nonetheless not related to prosody
given their insensitivity to stress.  More crucially, the tone 3
sandhi rule makes reference to the syntactic structure when it is
optional.  Therefore, it is considered a \enquote{syntactic} analysis
here.

The basic idea is that syntactic branching leads to different foot
structures.  For example, a left branching syntactic structure leads
to a left foot:

\ex.
\a. \(\text{(\textform{mǎi hǎo}) \textform{jiǔ}} \to
\text{(\textform{mái háo}) \textform{jiǔ}}\)
\bg. [[mǎi hǎo] jiǔ]\\
\phantom{[[}buy good wine\\
\trans\textgls{to finish buying wine}

Tone 3 sandhi applies to \textform{mǎi} as it is followed by another
tone 3 syllable in the same foot.  The curious part is \textform{hǎo},
which is followed by another tone 3 syllable in a \textemph{higher}
group.  A right branching syntactic structure, on the other hand,
leads to a right foot:

\ex.
\a. \(\text{\textform{mǎi} (\textform{hǎo jiǔ})} \to
\{\text{\textform{mǎi} (\textform{háo jiǔ})},
\text{\textform{mái} (\textform{háo jiǔ})}\}\)
\bg. [mǎi [hǎo jiǔ]]\\
\phantom{[}buy \phantom{[}good wine\\
\trans\textgls{to buy good wine}

Tone 3 sandhi now only optionally applies to \textform{mǎi}, as it is
followed by another tone 3 syllable in a \textemph{lower} group.
Apparently, this implies that nestable groupings are required.  As
aforementioned, nestable groupings lead to monadic second-order logic
over \textemph{trees}, which is an undesirable property.

Alternatively, we may consider that the trees be transformed to linear
sequences suitable as inputs to the later level.  This can be a simple
recursive process that flattens the input trees using an associative
operation that prefixes boundaries.\footnote{This process can also be
  represented as a strictly local tree transducer, given that linear
  sequences are isomorphic to single-pathed graphs, that is,
  unary-branching trees.  See \textcite{ioj20qtt, jh20isltt} for
  detailed discussions of such transducers.}  This tree transformation
can be nondeterministic to account for the optionality, but the level
it feeds, tone 3 sandhi, remains deterministic.\footnote{Under
  nondeterminism, an output sequence that contains a
  \(\check{\sigma}\check{\sigma}\) substring simply fails and
  backtracks.  Although \textcite{c00tspcd} reports several data that
  contain such substrings, they are more likely due to emphatic
  stresses.}  Considering only the nontrivial cases, we have

\ex.
\a. \(
\begin{forest}
  nice empty nodes,
  [[\(v\), baseline] [[\(w\)] [\(z\)]]]
\end{forest}
\to
\#v\#wz\)
\b. \(
\begin{forest}
  nice empty nodes,
  [[[\(v\)] [\(w\)]] [\(z\), baseline]]
\end{forest}
\to
\#\#vwz\)

In some sense, this is as if only opening brackets were preserved.
Syntactic categories are insignificant, as the process only makes
reference to the hierarchical structure.  This, of course, still
requires modification to tone 3 sandhi, as now the input sequences
contain boundaries.  To completely suppress boundaries in the inputs,
we can further split linear sequences by boundaries, resulting in
multiple linear sequences.\footnote{Or, obviously, transform to
  multiple linear sequences in a single pass.  The presentation is
  merely for easier understanding under a graph-transduction
  perspective.}  Tone 3 sandhi is then lifted to operate on these
linear sequences, whose outputs are combined by concatenation.  To put
it simply, this means that

\ex.
\a. \(\#v\#wz \to \tau(v) \cdot \tau(w \cdot z)\)
\b. \(\#\#vwz \to \tau(v \cdot w \cdot z)\)

omitting empty sequences.  Interestingly, the contrast now comes from
the fact that tone 3 sandhi does not preserve concatenation:

\ex. \(\neg(\forall x, y\ldotp
\tau(x \cdot y) = \tau(x) \cdot \tau(y))\)

If this \textemph{did} hold, we would be able to derive the
equivalence of the two mappings due to associativity.

A note should be made regarding cyclicity.  In \posscite{c00tspcd}
original analysis, cyclicity is used to explain both simultaneous and
optional applications, on the assumption that groupings can be
expanded at each level.  However, we have seen how cyclicity is
\textemph{not} an essential property of the analysis by virtue of
modularization, where a tree transformation feeds tone 3 sandhi.

\subsection{Prosodic Analysis}
\textcite{d07psc} proposes an alternative analysis in light of several
perceived inadequacies of the previous syntactic analysis.  The most
significant one is that supposedly same syntactic structures can lead
to different groupings, as

\ex. \(\text{(\textform{něizhǒng}) (\textform{jiǔ hǎo})} \to
\text{(\textform{néizhǒng}) (\textform{jiú hǎo})}\)

must be derived from

\exg. [[[něi-zhǒng] jiǔ] hǎo]\\
\phantom{[[[}which-kind alcoholic.beverage good\\
\trans\textgls{Which kind of alcoholic beverage is good?}

supposing the syntactic structure is correct at all.  This obviously
hinges on specific syntactic analysis, as one can imagine

\exg. [[[něi-zhǒng] [\gap{} jiǔ]] hǎo]\\
\phantom{[[[}which-kind \phantom{[}\textfeat{gap} alcoholic.beverage
good\\
\trans\textgls{Which kind of alcoholic beverage is good?}

being a totally valid alternative analysis that \textemph{does} derive
the correct grouping.\footnote{Justification for a similar analysis
  based on type shifting can be found in \textcite{bs14cnl}, which
  generalizes to all determiners.}  With the addition of gap terms,
the tree transformation requires more elaboration, but this does not
change the computational nature of the linear transformation.
Similarly, some other provided data can arguably be attributed to
alternative syntactic analyses.

Another perceived inadequacy is the insensitivity to stresses, which
is the reason behind the name \enquote{stress-insensitive foot}.  For
example, an emphatic stress as in

\ex. \(\text{\textform{xiǎng} (\textform{MǍI gǔpiào})} \to
\text{\textform{xiǎng} (\textform{MÁI gǔpiào})}\)

seems to favor (force?) a boundary before the emphasized syllable.
Again, this only requires modification to the tree transformation.
Nonetheless, \textcite{d07psc} opts for a more aggressive modification
to both the tree transformation and tone 3 sandhi.

Under the proposal, syntactic trees are assigned stresses according to
the \enquote{information-stress principle}, which covers syntactic
structures assuming that nonheads carry more information.  The
syllables are then grouped into feet based on stresses, where a
stressed syllable must starts a group.  As these feet are considered
metrical due to the relevance of stresses, the analysis is a prosodic
one.

The complication is that tone 3 sandhi must be aware of
\textemph{both} the syntactic structure and prosodic groupings due to
the way \enquote{adjacency} is defined.  Two syllables are said to be
\enquote{adjacent} when they are contained by the same syntactic
constituent and not by separate full feet, where a full foot is a
grouping with more than one syllable.  This notion must be defined
over autosegmental representations:

\ex.
\a. \(\mathit{Adjacent}(x, y) \coloneq \mathit{SameConst}(x, y) \land
\neg\mathit{SepFullFeet}(x, y)\)
\b. \(\mathit{SameConst}(x, y) \coloneq \exists z\ldotp
\mathit{Dom}(z, x) \land \mathit{Dom}(z, y)\)
\b. \(\mathit{SepFullFeet}(x, y) \coloneq \exists z, w\ldotp
z \neq w \land \mathit{FullFeet}(z, x) \land \mathit{FullFeet}(w, y)\)
\b. \(\mathit{FullFeet}(x, y) \coloneq \mathit{Feet}(x, y) \land
(\exists z\ldotp \mathit{Feet}(x, z) \land
(\mathit{prec}(y) = z \lor \mathit{succ}(y) = z))\)

\(\mathit{Dom}\) is the transitive closure of the syntactic
association relation, and \(\mathit{Feet}\) is the prosodic
association relation.  This is overcomplicated as a phonological
process, at least from our point of view.

Moreover, tone 3 sandhi is now said to be a nondeterministic process,
which optionally changes tone 3 before nonadjacent or sandhi tone 3.
In a logical world, this can be achieved by an operator that
introduces choice points, say \posscite{m61bmtcpr} \(\mathit{amb}\)
operator:

\ex.
\a. \(\acute{\sigma}'(x) \coloneq \acute{\sigma}(x) \lor
(\check{\sigma}(x) \land \check{\sigma}(\mathit{succ}(x))
\land \neg\mathit{MaybeSkip}(x))\)
\b. \(\mathit{MaybeSkip}(x) \coloneq
\mathit{amb}(\neg\mathit{Adjacent}(x, \mathit{succ}(x)),
\check{\sigma}(\mathit{succ}(\mathit{succ}(x))), \bot)\)

We see how convoluted the analysis becomes once the phonological
process is no longer subregular.  Even though tone 3 sandhi is a one
of the simplest phonological processes, the \textemph{condition} it
applies heavily depends on syntactic information.  Still worse, its
application can be nondeterministic.  By distributing the workload to
tone 3 sandhi, its computational nature is effectively obscured.
Moreover, using autosegmental representations to represent
\textemph{hierarchical} structures is in some sense an abuse.

\section{Tianjin Mandarin: Rule Interaction}
Tianjin Mandarin is a more interesting case than Standard Chinese, as
there are a total of \textemph{three} tone sandhi rules that interact
in a nontrivial way.  In Tianjin Mandarin, three dissimilation rules
similar to tone 3 sandhi in Standard Chinese are found:\footnote{For
  convenience, notations borrowed from Hanyu Pinyin are used to
  indicate the tones, although they do not phonologically correspond
  to the specific tones in Standard Chinese.}

\ex.
\a. \(\grave{\sigma}\grave{\sigma} \to \check{\sigma}\grave{\sigma}\)
\hfill(F[alling-]F sandhi)
\b. \(\check{\sigma}\check{\sigma} \to \acute{\sigma}\check{\sigma}\)
\hfill(L[ow-]L sandhi)
\b. \(\acute{\sigma}\acute{\sigma} \to \bar{\sigma}\acute{\sigma}\)
\hfill(R[ising-]R sandhi)

These rules approximately generalize to trisyllabic sequences in a
manner reminiscent of directionalities:

\ex.
\a. \(\grave{\sigma}\grave{\sigma}\grave{\sigma} \to
\grave{\sigma}\check{\sigma}\grave{\sigma}\)
\hfill(right-to-left)
\b. \(\check{\sigma}\check{\sigma}\check{\sigma} \to
\check{\sigma}\acute{\sigma}\check{\sigma}\)
\hfill(right-to-left)
\b. \(\acute{\sigma}\acute{\sigma}\acute{\sigma} \to
\bar{\sigma}\bar{\sigma}\acute{\sigma}\)
\hfill(left-to-right)

There has been debates on how these rules interact and, by extension,
how to explain such interactions.  Rule-based analyses such as
\posscite{t87tstd, z87ptal} proposals rely on stipulations to account
for the directionalities as well as the feeding relations, therefore
offering no real explanation.  In the subregular zoo, instead of
directionalities, the difference in patterns amounts to the difference
in logical formulae, that is, the computational nature of the rules.
The problem is then how to also accommodate the feeding relations.  A
relevant \enquote{boundaryless} computational analysis will be
discussed before a new \enquote{boundaryful} analysis is presented.

\subsection{Bounadaryless Analysis}
\posscite{c18catsi} analysis is based on the observation that the
three strictly local functions can be combined into a single one,
which must remember a pair of input and output substrings.  It is thus
an input-output strictly local function.  Following our convention so
far, the idea will be described using logical formulae instead of
transducers.  The logical transduction, without consideration of the
feeding relations, is as follows:

\ex.
\a. \(\check{\sigma}'(x) \coloneq
(\check{\sigma}(x) \land \neg\check{\sigma}'(\mathit{succ}(x))) \lor
(\grave{\sigma}(x) \land \grave{\sigma}'(\mathit{succ}(x)))\)
\b. \(\acute{\sigma}'(x) \coloneq
(\acute{\sigma}(x) \land \neg\acute{\sigma}(\mathit{succ}(x))) \lor
(\check{\sigma}(x) \land \check{\sigma}'(\mathit{succ}(x)))\)
\b. \(\bar{\sigma}'(x) \coloneq \bar{\sigma}(x) \lor
(\acute{\sigma}(x) \land \acute{\sigma}(\mathit{succ}(x)))\)

In the transduction, rule interactions are naturally derived from the
shapes of logical formulae.  We see that FF sandhi feeds LL sandhi
given the presence of \(\check{\sigma}'\), and similarly that LL
sandhi counterfeeds RR sandhi given the absence of
\(\acute{\sigma}'\).  A redefinition of the transduction gives us
another possibility:

\ex.
\a. \(\check{\sigma}'(x) \coloneq
(\check{\sigma}(x) \land
\neg\check{\sigma}^{\land}(\mathit{succ}(x))) \lor
(\grave{\sigma}(x) \land \grave{\sigma}'(\mathit{succ}(x)))\)
\b. \(\acute{\sigma}'(x) \coloneq
(\acute{\sigma}(x) \land
\neg\acute{\sigma}^{\lor}(\mathit{succ}(x))) \lor
(\check{\sigma}(x) \land \check{\sigma}^{\land}(\mathit{succ}(x)))\)
\b. \(\bar{\sigma}'(x) \coloneq \bar{\sigma}(x) \lor
(\acute{\sigma}(x) \land
\acute{\sigma}^{\lor}(\mathit{succ}(x)))\)
\b. \(\check{\sigma}^{\land}(x) \coloneq
\check{\sigma}(x) \land \check{\sigma}'(x)\)
\b. \(\acute{\sigma}^{\lor}(x) \coloneq
\acute{\sigma}(x) \lor \acute{\sigma}'(x)\)

FF sandhi now counterfeeds LL sandhi given the presence of
\(\check{\sigma}^{\land}\), that is, a sandhi low tone does
\textemph{not} trigger LL sandhi.  Likewise, LL sandhi now feeds RR
sandhi given the presence of \(\acute{\sigma}^{\lor}\), that is, a
rising tone either in the input or output sequence triggers RR sandhi.

This analysis apparently does not make reference to boundaries, unlike
our analysis of tone 3 sandhi in Standard Chinese.  This leads a
difficulty in describing the conflicting data:

\ex. \(\check{\sigma}\grave{\sigma}\grave{\sigma} \to
\{\check{\sigma}\check{\sigma}\grave{\sigma},
\acute{\sigma}\check{\sigma}\grave{\sigma}\}\)

The former alternative suggests that FF sandhi counterfeeds LL sandhi,
while the latter suggests otherwise:

\ex.
\a. \(\check{\sigma}\grave{\sigma}\grave{\sigma} \to
\check{\sigma}\check{\sigma}\grave{\sigma}\)
\b. \(\check{\sigma}\grave{\sigma}\grave{\sigma} \to
\check{\sigma}\check{\sigma}\grave{\sigma} \to
\acute{\sigma}\check{\sigma}\grave{\sigma}\)

The presence of the extra step of application is exactly what
differentiates counterfeeding and feeding, at least given the lack of
boundaries.  Unfortunately, we can encode either counterfeeding or
feeding, but not both.  The contrast of the two, nonetheless, is
analogous to the situation in Standard Chinese.  Indeed, the following
holds assuming FF sandhi feeds LL sandhi:

\ex.
\a. \(\tau(\check{\sigma}) \cdot \tau(\grave{\sigma}\grave{\sigma}) =
\check{\sigma}\check{\sigma}\grave{\sigma}\)
\b. \(\tau(\check{\sigma}\grave{\sigma}\grave{\sigma}) =
\acute{\sigma}\check{\sigma}\grave{\sigma}\)

\subsection{Boundaryful Analysis}
The data shown in \textcite{t87tstd} are actually subject to a clear
contrast between left-branching and right-branching structures, which
we now know derive different boundaries.  The most trivial case of
boundaries is found when every syllable is emphatically stressed, thus
creating the sequence

\ex. \(v\#w\#z\)

that does not undergo any sandhi.  This happens since a single
syllable can not be affected by the transduction, namely that

\ex. \(\tau(v) \cdot \tau(w) \cdot \tau(z) = vwz\)

This is not particularly interesting, but does show that boundaries
are in effect.  This case is henceforth ignored for lack of interest.

The more interesting cases are found when some syllables do undergo
sandhi.  In these cases, Tianjin Mandarin uses a more
\enquote{conservative} approach to creating boundaries, such that
every constituent receives a prefix and suffix boundary when
flattened.  This leads to the tree transformation

\ex.
\a. \(
\begin{forest}
  nice empty nodes,
  [[\(v\), baseline] [[\(w\)] [\(z\)]]]
\end{forest}
\to
\#v\#wz\#\#\)
\b. \(
\begin{forest}
  nice empty nodes,
  [[[\(v\)] [\(w\)]] [\(z\), baseline]]
\end{forest}
\to
\#\#vw\#z\#\)

that appears as if every bracket is rewritten into a boundary term.
In other words, a left-branching structure causes the substring of
first two syllables to undergo tone sandhi, while a right-branching
structure causes the substring of last two syllables to undergo tone
sanhi, when boundaries are outputted at all.

As usual, the tree transformation is nondeterministic, so that
boundaries are allowed to be omitted.  With this in mind, it is still
true that the patterns that are attested for \textemph{both}
left-branching and right-branching structures can be used to justify
the trisyllabic generalization, as these patterns occur when
boundaries are omitted.  The transduction therefore remains strictly
local, while the originally conflicting data can now be well
explained:

\ex. \(\check{\sigma}(\grave{\sigma}\grave{\sigma}) \to
\{\check{\sigma}\#\check{\sigma}\grave{\sigma},
\acute{\sigma}\check{\sigma}\grave{\sigma}\}\)

Applying this idea to more data shows that it is indeed the right
track.  The cases of trisyllabic sequences are shown first, as these
cases do not involve the complication of feeding relations:

\ex.
\a. \(\check{\sigma}(\check{\sigma}\check{\sigma}) \to
\{\check{\sigma}\#\acute{\sigma}\check{\sigma},
\check{\sigma}\acute{\sigma}\check{\sigma}\}\)
\b. \((\check{\sigma}\check{\sigma})\check{\sigma} \to
\{\acute{\sigma}\check{\sigma}\#\check{\sigma},
\check{\sigma}\acute{\sigma}\check{\sigma}\}\)
\b. \(\acute{\sigma}(\acute{\sigma}\acute{\sigma}) \to
\{\acute{\sigma}\#\bar{\sigma}\acute{\sigma},
\bar{\sigma}\bar{\sigma}\acute{\sigma}\}\)
\b. \((\acute{\sigma}\acute{\sigma})\acute{\sigma} \to
\{\bar{\sigma}\acute{\sigma}\#\acute{\sigma},
\bar{\sigma}\bar{\sigma}\acute{\sigma}\}\)

The FF sandhi data also suggest a minor modification to the
transduction to accommodate the rule

\ex. \(\grave{\sigma}\check{\sigma} \to \bar{\sigma}\check{\sigma}\)
\hfill(FL sandhi)

that produces the following mappings:

\ex.
\a. \(\grave{\sigma}\grave{\sigma}\grave{\sigma} \to
\bar{\sigma}\check{\sigma}\grave{\sigma}\)
\b. \(\grave{\sigma}\grave{\sigma}\check{\sigma} \to
\check{\sigma}\bar{\sigma}\check{\sigma}\)

This can be done by redefining the logical transduction:

\ex.
\a. \(\check{\sigma}'(x) \coloneq
(\check{\sigma}(x) \land
\neg\check{\sigma}^{\land}(\mathit{succ}(x))) \lor
(\grave{\sigma}(x) \land
\grave{\sigma}(\mathit{succ}(x)) \land
\neg\check{\sigma}'(\mathit{succ}(x)))\)
\b. \(\bar{\sigma}'(x) \coloneq \bar{\sigma}(x) \lor
(\acute{\sigma}(x) \land
\acute{\sigma}^{\lor}(\mathit{succ}(x))) \lor
(\grave{\sigma}(x) \land
\check{\sigma}'(\mathit{succ}(x)))\)

The interaction is a curious one, as now FF sandhi feeds FL sandhi
\textemph{while} counterbled by it.  This is rather complicated to
describe with cyclicity, but simple enough as a strictly local
function.  Correct patterns of FF sandhi can now be predicted:

\ex.
\a. \(\grave{\sigma}(\grave{\sigma}\grave{\sigma}) \to
\{\grave{\sigma}\#\check{\sigma}\grave{\sigma},
\bar{\sigma}\check{\sigma}\grave{\sigma}\}\)
\b. \((\grave{\sigma}\grave{\sigma})\grave{\sigma} \to
\{\check{\sigma}\grave{\sigma}\#\grave{\sigma},
\bar{\sigma}\check{\sigma}\grave{\sigma}\}\)

To decide the feeding relations, data involving rule interactions are
examined.  LL sandhi bleeds FL sandhi:

\ex.
\a. \(\grave{\sigma}(\check{\sigma}\check{\sigma}) \to
\{\grave{\sigma}\#\acute{\sigma}\check{\sigma},
\grave{\sigma}\acute{\sigma}\check{\sigma}\}\)
\b. \((\grave{\sigma}\check{\sigma})\check{\sigma} \to
\{\bar{\sigma}\check{\sigma}\#\check{\sigma},
\grave{\sigma}\acute{\sigma}\check{\sigma}\}\)

FL sandhi counterbleeds FF sandhi:

\ex.
\a. \(\grave{\sigma}(\grave{\sigma}\check{\sigma}) \to
\{\grave{\sigma}\#\bar{\sigma}\check{\sigma},
\check{\sigma}\bar{\sigma}\check{\sigma}\}\)
\b. \((\grave{\sigma}\grave{\sigma})\check{\sigma} \to
\{\check{\sigma}\grave{\sigma}\#\check{\sigma},
\check{\sigma}\bar{\sigma}\check{\sigma}\}\)

LL sandhi feeds RR sandhi:

\ex.
\a. \(\acute{\sigma}(\check{\sigma}\check{\sigma}) \to
\{\acute{\sigma}\#\acute{\sigma}\check{\sigma},
\bar{\sigma}\acute{\sigma}\check{\sigma}\}\)
\b. \((\acute{\sigma}\check{\sigma})\check{\sigma} \to
\{\acute{\sigma}\check{\sigma}\#\check{\sigma},
\bar{\sigma}\acute{\sigma}\check{\sigma}\}\)

FF sandhi feeds LL sandhi:

\ex.
\a. \(\check{\sigma}(\grave{\sigma}\grave{\sigma}) \to
\{\check{\sigma}\#\check{\sigma}\grave{\sigma},
\acute{\sigma}\check{\sigma}\grave{\sigma}\}\)
\b. \((\check{\sigma}\grave{\sigma})\grave{\sigma} \to
\{\check{\sigma}\grave{\sigma}\#\grave{\sigma},
\acute{\sigma}\check{\sigma}\grave{\sigma}\}\)

All that remains is a transduction that is compatible with all cases,
which can be defined using the shown techniques.

Lastly, some rare cases in the data seem to suggest that LL sandhi
produces the alternative mapping:

\ex. \(\check{\sigma}\check{\sigma}\check{\sigma} \to
\acute{\sigma}\acute{\sigma}\check{\sigma}\)

As previously discussed, adding alternatives to the transduction
incurs nondeterminism.  Instead, it is more likely that some lexical
items with a low tone have alternative realizations with a rising
tone, moving nondeterminism out from the transduction.\footnote{This
  can also explain the mapping
  \(\check{\sigma}\grave{\sigma}\grave{\sigma} \to
  \acute{\sigma}\check{\sigma}\grave{\sigma}\) even if counterfeeding
  is assumed.}

\section{Interim Summary}
This section has argued for a modular analysis of tonal transformation
that is sensitive to the hierarchical structure.  The point is not to
suggest that a certain analysis is inherently better than another.
Rather, it argues that such a modular analysis is computationally
feasible and clarifies the boundaries between linguistic levels.  Even
if the overall complexity of the composed system remains the same, a
modular analysis allows for the local reasoning of each module due to
abstraction and encapsulation.  This is also why modularization is
heavily studied in programming language theory.

Specifically, a boundary-based analysis is presented for both Standard
Chinese and Tianjin Mandarin, where a boundary is merely a term that
blocks the transformation.  In this sense, the whole transformation
can be understood as a boundary-agnostic subtransformation
\enquote{lifted} to operate on multiple linear sequences that are
produced by splitting the input sequence by boundaries.  This is how we
achieve modularity---by further and further separating out each step
of computation and composing them together at the end.

The theoretical implication is that the methodological assumption of
lexical phonology or any modular approach in general is preferable.
Nonetheless, this does not and should not bear any substantive
implication.  Nothing is assumed about the \enquote{architecture of
  the language faculty}, as this issue is simply not important in the
computational understanding of phonology.  We agree with lexical
phonology insomuch as both are opposed to the cyclic and thus
monolithic view of phonology.  Anything more is left aside.

\chapter{Power of Representation}
In the previous chapter, we have had a taste of how representations
can obscure the computational nature.  This chapter relates this
intuition to even more attested tonal transformations.
Representations are indispensable both as a theoretical tool and
metatheoretical notion, and they often enforce specific ways to
express the computation.  Choosing an incorrect representation can
either overly simplify or complicate the computation, much to our
dismay.

\section{Representation as Computation}
Representations, when considered as a part of bigger computation,
implicitly \textterm{encode} partially or fully done computation.
This is not new to linguists, as syntacticians often face the
dichotomy between features and constraints, whose equivalence is
pointed out in \textcite{g17cgdfc}.  The contentious issue of enriched
representations has also been a long-standing problem in theoretical
phonology.
